Query Optimization for Dynamic Imputation

### 2017-04-25 rebuttal.tex

\documentclass[draft,12pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{xcolor}
\newcommand{\resp}[1]{{\color{blue}{#1}}}

@@

\item - S1: "Furthermore, the results obtained with this approach are similar in quality to those obtained with the traditional imputation approach — between 0 and 24 percent" 24 percent error is pretty bad.

\item - S2: The drop operator appears to be a weaker form of SQL's null-value / 3-valued Boolean semantics. The latter might actually be more useful. For example, consider the case of a predicate (A $>$ 5 OR B $<$ 10). If A $>$ 5, then the condition is independent of B, and the tuple can still be salvaged. If A $\leq$ 5, then the expression evaluates to UNKNOWN and the tuple is dropped anyway. For aggregates there is a greater need, as one might have different aggregates computed over different subsets of the data, but even here it's not clear that this is any less informative than just throwing data away.

\item - S2: Why no unions?

\resp{We skipped unions for brevity. There is nothing fundamental about our algorithm which makes unions difficult to implement.}

@@

\item - S3.4.1, Definition 2: On my first read-through, I was puzzled about why there was no plan with zero-loss. It would help to clarify that Loss is not about tuple loss, but rather a heuristic penalty function.

\item - S3.4.3: An extremely nice emergent property from the L-value function for the imputation operator seems to be that the operator will be moved to a position in the plan where it has access to the maximum training data. I am slightly concerned that by focusing entirely on size, this might lead to skew in the training set. For example, consider a one-to-many join \verb/µ(R) |X| S/. Your optimization algorithm would likely wish to move imputation up: \verb/µ(R |X| S)/. Yes, there is more training data, but if the fanout for each tuple is skewed, some tuples in R will become over-represented in the join result. A similar observation holds for selection. For example take your Q8, where the conditions "studentdebtowe $>$ 0 AND schooldegree $\geq$ 0" presumably indicate invalid data that should not be used for training. (is this perhaps why you have so much error for full imputation on Q8?)

  % \resp{This is a reasonable concern. Optimizing for the maximum training set is flawed when the training data is skewed. We think that this problem is avoidable with a minor tweak to the loss function. Instead of optimizing for a large training set, we can optimize for a \emph{diverse} training set by tracking the

@@

\item - S3.4.4: It appears that Time() assumes that plans are IO-bound. This is a perfectly reasonable assumption, but should be stated explicitly, as datasets that fit in memory will require a different cost model.

  \resp{Agreed. Our database stores tables in files, so our plans are I/O bound, but a database which uses an in-memory representation will need a different Time() function.}

@@

\item - There are a number of opportunities for trimming in the figures, tables, and algorithms: Algorithms 1-3 could be compacted a bunch, Table 1 is full of whitespace (e.g., make it a 3-column {figure*}), Tables 1-2 could also be summarized as CDFs, Figure 6 is full of whitespace, Table 3 could be put queries on a single row (e.g., Q1 would easily fit on 1 line), and the references have a lot of redundant information (URLs, for example).

\item - Additional space recovered from these sections could be used to discuss the choice of techniques for imputation, and the confidence estimation idea suggested in the future work section.

\item - Another interesting use of the space, or at least an idea for future work might be the role of aggregation on inference accuracy. In my experience aggregate queries tend to be *much* more forgiving of approximations (e.g., Online Aggregation) and data errors.
\end{itemize}

\section{Reviewer 2}
\begin{itemize}
\item 1) No new imputation method is presented. The model relies on external imputation techniques, and given it's done on-the-fly, presumably imputed values depend on qualifying query records.

  \resp{Yes, the imputed results depend on the records which are relevant to particular query. We do not present a new imputation method, but our technique is agnostic to the particular imputation method used. We reduce the overhead of imputation by performing it only on the tuples which are relevant to a particular query.}

\item 2) No comparative evaluation against existing imputation/data cleaning techniques.

  \resp{Our evaluation compares our technique to an imputation of the full dataset. We use the same imputation algorithm for our on-the-fly imputation as for the full dataset imputation. Cleaning the full dataset is the standard method of handling missing values.}

@@

\item D1) The model relies on external imputation techniques, and given it's done on-the-fly, the imputed values will differ based on the qualifying query records?

  \resp{Yes, as discussed above.}

\item D2) The paper does not propose new imputation techniques but rely on existing statistical imputation algorithms. There is no comparative evaluation, nor samples of imputed values to show the goodness of the imputed values. It would be nice to see a comparative evaluation of different statistical imputation methods, and suggestions for extensions that can fit within the proposed imputation model.

  \resp{The error measures that we provide compare our on-the-fly application of imputation to imputation over the full dataset.}

@@

\item W3. More experiments needed.

\item D1. Section 3, which is the meat of the paper, is very overloaded, without getting in depth in various parts of the system. It definitely should be split in more sections, and more details should be given. In particular:
-- Give some more information about cardinality estimation. Also, Algorithm 3 is too far from the text, and needs some explanation too.
-- Some more details are also required in the Imputation quality section. Why are you choosing these specific formulas for Loss, especially for the impute operator. Some examples would help.
-- Algorithms 1 and 2 could probably be presented in an easier to grasp way.

\item D2. It would be nice to show how exactly your approach can be implemented in an existing optimizer. This comes up in various places in the paper:
-- When defining the search space, why not first define the transformation operators that inject the imputation operators. Then any optimizer can pick these rules up and add them in its plan enumeration. Of course, the used approach that follows System-R's plan enumeration is useful, but why lose generality when presenting the search space. Moreover, showing the transformation rules can help you show the size of the search space that will better justify the restrictions you apply.
-- Why use simplified cost formulas for the traditional relational operators and not existing systems' formulas?
-- It would be great (but not necessary of course) to add your techniques in an existing DBMS, such as Postgres to see the impact on a real system in your experiments.

\resp{Existing systems have very complex formulas for estimating query costs. We felt that the overhead of reimplementing an existing cost model would not be worth the benefit.}

@@

\item D3. Is there any way for the user to know what error to expect based on the $\alpha$ value? Moreover, is its value somehow proportional to the error (e.g., $\alpha=0.5$ would give roughly double the error of $\alpha=0.25$)? Or should the user simply try multiple execution with different values of $\alpha$ until the desired results are given?

  \resp{In our intended usage, the user starts by running with $\alpha = 0$ and increases $\alpha$ if the query takes too long to complete.}

@@

\item D4. How would a two-step approach that first finds the optimal plan (w/o imputation) and then adds the imputation operators perform?

  \resp{This approach would not be able to improve on the plans produced by our approach, as it would produce a subset of the plans that we consider. It's unclear how much worse the plans would be, but a major issue is that the Drop and Impute operators have an impact on the cardinality of their outputs. Our optimizer can use this information to improve the plan, but adding imputation after the fact would not be able to rearrange the plan to capitalize on changes in cardinality.}

@@

\item D5. Can you have plans with a mix of drop and impute operators?

  \resp{Yes.}

\item D6. During plan enumeration, can you safely prune plans at some point, similar to what a traditional optimizer would do?

  \resp{We maintain the set of pareto optimal plans at each point in the optimization. This set is significantly smaller than the set of all plans, so it represents some pruning. We found that the size of these pareto sets is small enough that we did not need to prune further.

    If necessary, it should be possible to sample from the pareto optimal plans during the plan enumeration. The final selected plan may not be pareto optimal, but this is a potential way to control the number of plans.}

@@

\item D8. Are there impute implementation that are non-blocking? Or do they always have to perform multiple passes over the data?

\item D9. Additional experiments I suggest to include:
-- Consider a scenario where the user submits multiple similar queries. Show after how many such queries your approach has no longer a competitive advantage compared to performing imputation on the data.
-- Try different values for $\alpha$ (not just 0 and 1).
-- Show runtime of enumeration algorithm as queries get more complex.

\item D10. The scenario you are considering, with multiple queries during exploration, begs for reuse of previous query results, as you also mention in the future work section.

\item D11. Less important points:
-- Sections 4.2 and 4.3 in the experiments are too verbose.
-- In Section 3.2, you mention that joins are performed after filtering. When filters are on results of joins, this is not possible.

\item -- Given you optimize for both performance and quality, Tummer's and Koch's work on multi-objective QO is related.

  \resp{Yes, Tummer \& Koch is closely related. However, much of the complexity of their paper is in dealing with cost functions which are not well behaved. Our cost functions are both monotonic, which allows us to use a simpler optimizer than what Tummer \& Koch propose.}

\item -- Could be interesting for the optimizer to pick imputation strategy based on the type of the query.
\end{itemize}
\end{document}
%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2017-04-25 paper/main.tex
paper: rework queries table to make it smaller

% listings settings
\lstset{ %
  backgroundcolor=\color{white},   % choose the background color; you must add \usepackage{color} or \usepackage{xcolor}; should come as last argument
  basicstyle=\ttfamily\footnotesize,        % the size of the fonts that are used for the code
  breakatwhitespace=false,         % sets if automatic breaks should only happen at whitespace
  breaklines=true,                 % sets automatic line breaking
  captionpos=b,                    % sets the caption-position to bottom
  commentstyle=\color{red},    % comment style
  %frame=single,	                   % adds a frame around the code
  keepspaces=true,                 % keeps spaces in text, useful for keeping indentation of code (possibly needs columns=flexible)
  keywordstyle=\color{blue},       % keyword style
%  numbers=left,                    % where to put the line-numbers; possible values are (none, left, right)
  numbersep=5pt,                   % how far the line-numbers are from the code
%  numberstyle=\tiny\color{gray}, % the style that is used for the line-numbers
  rulecolor=\color{black},         % if not set, the frame-color may be changed on line-breaks within not-black text (e.g. comments (green here))
%  stepnumber=2,                    % the step between two line-numbers. If it's 1, each line will be numbered
  stringstyle=\color{red},     % string literal style
  tabsize=2,	                   % sets default tabsize to 2 spaces
  language=SQL
}

>>>>

% listings settings
\lstset{ %
  backgroundcolor=\color{white},   % choose the background color; you must add \usepackage{color} or \usepackage{xcolor}; should come as last argument
  basicstyle=\ttfamily\footnotesize,        % the size of the fonts that are used for the code
  breakatwhitespace=true,         % sets if automatic breaks should only happen at whitespace
  breaklines=true,                 % sets automatic line breaking
  captionpos=b,                    % sets the caption-position to bottom
  commentstyle=\color{red},    % comment style
  %frame=single,	                   % adds a frame around the code
  keepspaces=true,                 % keeps spaces in text, useful for keeping indentation of code (possibly needs columns=flexible)
  keywordstyle=\color{blue},       % keyword style
%  numbers=left,                    % where to put the line-numbers; possible values are (none, left, right)
  numbersep=5pt,                   % how far the line-numbers are from the code
%  numberstyle=\tiny\color{gray}, % the style that is used for the line-numbers
  rulecolor=\color{black},         % if not set, the frame-color may be changed on line-breaks within not-black text (e.g. comments (green here))
%  stepnumber=2,                    % the step between two line-numbers. If it's 1, each line will be numbered
  stringstyle=\color{red},     % string literal style
  tabsize=2,	                   % sets default tabsize to 2 spaces
  language=SQL
}

### 2017-04-25 paper/queries.tex
paper: rework queries table to make it smaller

\newcounter{queryno}
\begin{tabular}{cl}
\toprule
\# & \multicolumn{1}{c}{Query} \\
\midrule
1 & \verb|SELECT BLD FROM acs;| \refstepcounter{queryno} \label{q1} \\
2 & \verb|SELECT BLD as units_in_structure, COUNT(ST) as estimate FROM acs GROUP BY BLD;| \refstepcounter{queryno} \label{q2} \\
3 & \verb|SELECT BATH as has_bath, COUNT(ST) as ct FROM acs GROUP BY BATH;| \refstepcounter{queryno} \label{q3} \\
4 & \verb|SELECT ACR as lotsize, AVG(BDSP) as avg_num_bedrooms FROM acs GROUP BY ACR;| \refstepcounter{queryno} \label{q4} \\
5 & \verb|SELECT PSF as has_sub_families, SUM(NP) as num_people FROM acs GROUP BY PSF;| \refstepcounter{queryno} \label{q5} \\
6 & \verb|SELECT AVG(NP) as avg_num_people FROM acs;| \refstepcounter{queryno} \label{q6} \\
7 & \verb|SELECT BDSP as num_bedrooms, AVG(ACR) as avg_lot_size FROM acs WHERE VEH >= 2 GROUP BY BDSP;| \refstepcounter{queryno} \label{q7} \\
8 & \verb|SELECT MIN(RMSP) as min_num_rooms FROM acs WHERE RWAT=2;| \refstepcounter{queryno} \label{q8} \\
9 & \verb|SELECT * FROM dirty m where m.f1 >= 2;| \refstepcounter{queryno} \label{q9} \\
10 & \verb|SELECT * FROM dirty where dirty.f1 <= 2;| \refstepcounter{queryno} \label{q10} \\
11 & \verb|SELECT MAX(f2) as max_f2 FROM dirty GROUP BY dirty.f1;| \refstepcounter{queryno} \label{q11} \\
12 & \verb|SELECT m1.f1, m1.f2 FROM dirty m1, clean m2 WHERE m1.f1 = m2.f1 and m1.f1 = 1 and m2.f2 = 69;| \refstepcounter{queryno} \label{q12} \\
13 & \verb|SELECT d1.f1 FROM dirty d1, clean d2 WHERE d1.f1 = d2.f1 and d2.f2 = 50;| \refstepcounter{queryno} \label{q13} \\
\bottomrule
\end{tabular}

>>>>

\newcounter{queryno}
\begin{tabular}{cl}
\toprule
\# & \multicolumn{1}{c}{Queries on CDC data} \\
\midrule
1 & 
\begin{minipage}{6in}
\begin{lstlisting}[breaklines]
SELECT income, AVG(cuff_size) FROM demo, exams 
WHERE demo.id = exams.id AND height >= 150 GROUP BY income;
\end{lstlisting}
\end{minipage}\refstepcounter{queryno}\label[query]{q1} \\
2 & 
\begin{minipage}{6in}
\begin{lstlisting}[breaklines]
SELECT income, AVG(creatine) FROM demo, exams, labs 
WHERE demo.id = exams.id AND exams.id = labs.id AND income >= 13 AND income <= 15 AND weight >= 63 GROUP BY income;
\end{lstlisting}
\end{minipage}
\refstepcounter{queryno}\label[query]{q2} \\
3 & 
\begin{minipage}{6in}
\begin{lstlisting}[breaklines]
SELECT AVG(blood_lead) FROM demo, exams, labs 
WHERE demo.id = labs.id AND labs.id = exams.id AND age_yrs <= 6;
\end{lstlisting}
\end{minipage}\refstepcounter{queryno}\label[query]{q3}\\
4 & 
\begin{minipage}{6in}
\begin{lstlisting}[breaklines]
SELECT gender, AVG(blood_pressure_systolic) FROM demo, labs, exams 
WHERE demo.id = labs.id AND labs.id = exams.id AND body_mass_index >= 30 GROUP BY gender;
\end{lstlisting}
\end{minipage}\refstepcounter{queryno}\label[query]{q4}\\
5 & 
\begin{minipage}{6in}
\begin{lstlisting}[breaklines]
SELECT AVG(waist_circumference) FROM demo, exams 
WHERE demo.id = exams.id AND height >= 150 AND weight >= 100;
\end{lstlisting}
\end{minipage}\refstepcounter{queryno}\label[query]{q5}\\
\# & \multicolumn{1}{c}{Queries on freeCodeCamp data} \\
\midrule
6 & 
\begin{minipage}{6in}
\begin{lstlisting}[breaklines]
SELECT attendedbootcamp, AVG(income) FROM fcc
WHERE income >= 50000 GROUP BY attendedbootcamp;
\end{lstlisting}
\end{minipage}\refstepcounter{queryno} \label[query]{q6} \\
7 & 
\begin{minipage}{6in}
\begin{lstlisting}[breaklines]
SELECT AVG(commutetime) FROM fcc
WHERE gender = "female" AND countrycitizen = "United States";
\end{lstlisting}
\end{minipage}\refstepcounter{queryno} \label[query]{q7} \\
8 & 
\begin{minipage}{6in}
\begin{lstlisting}[breaklines]
SELECT schooldegree, AVG(studentdebtowe) FROM fcc
WHERE studentdebtowe > 0 AND schooldegree >= 0 GROUP BY schooldegree;
\end{lstlisting}
\end{minipage}\refstepcounter{queryno} \label[query]{q8}\\
9 & 
\begin{minipage}{6in}
\begin{lstlisting}[breaklines]
SELECT attendedbootcamp, AVG(gdp_per_capita) FROM fcc, gdp
WHERE fcc.countrycitizen = gdp.country AND age >= 18
GROUP BY attendedbootcamp;
\end{lstlisting}
\end{minipage}\refstepcounter{queryno} \label[query]{q9}\\
\bottomrule
\end{tabular}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End:
